---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---


# load packages
```{r}
rm(list = ls())
library(data.table)
library(reshape)
library(ggplot2)
library(scico)
library(png)

today = Sys.Date()
```

# load data
```{r}
# simulation parameters
#params = read.table("../config/parameters.tsv", sep = "\t", header = T)
params = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/selection-demography-cnn/config/parameters.tsv", sep = "\t", header = T)

# Ne values
#mean_ne = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/selection-demography-cnn/workflow/harmonic_mean_ne.txt", sep = " ", header = F, col.names = c("ID", "Ne"))
#params = merge(params, mean_ne, by = "ID")

# merge in fixation times
fix_times = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/selection-demography-cnn/workflow/fixation_times.txt", col.names = c("ID", "tf"))

params = merge(params, fix_times, by = "ID")

# plot fixation times for different types of sweeps
params$sweepType = NA
params$sweepType[( (params$f0 == 0) & (params$f1 == 1) )] = "hard"
params$sweepType[( (params$f0 > 0) & (params$f1 == 1) )] = "soft"
params$sweepType[( (params$f0 == 0) & (params$f1 < 1) )] = "partial"
params$sweepType[( (params$f0 > 0) & (params$f1 < 1) )] = "soft+partial"
table(params$sweepType)

ggplot(params, aes(x = log10(tf), color = sweepType)) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(Fixation time)", color = "Type")

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/selection-demography-cnn/results/", today, "/tf_distribution_by_sweep_type_", today, ".png", sep = ""), height = 7, width = 7)

# get random sample of simulation outputs
mypics = sample(list.files("/mnt/scratch/robe1195/Josephs_Lab_Projects/selection-demography-cnn/workflow/data/images", full.names = T), size = 1000, replace = F)

# get sim ids
picids = as.integer(gsub(".png", "", gsub("/mnt/scratch/robe1195/Josephs_Lab_Projects/selection-demography-cnn/workflow/data/images/slim_", "", mypics)))

# get variable that outputs should predict 
y = log10(params[(params$ID %in% picids), "tf"])
summary(y)

# define function for reading pngs,
read_png = function(x){
  print(x)
  temp_pic = readPNG(x)[,,1]
  melt(temp_pic)$value
}

# load sim outputs
mypics = lapply(mypics, read_png)

# bind all sim outputs together
mypics = do.call("cbind", mypics)

# convert gray pixels to 0.5
mypics[((mypics > 0) & (mypics < 1))] = 0.5

# simulations should be rows
# do PCA
pca_results = as.data.frame(prcomp(t(mypics), center = T)$x)

# visually show how PCs correlate with dependent variable
ggplot(pca_results, aes(PC1, PC2, color = y)) +
  geom_point() +
  theme_classic() +
  labs(title = "PCA of 5000 images") + 
  scale_color_scico(palette = "lajolla", name = "log10(Scaled fixation time)") 

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/selection-demography-cnn/results/", today, "/pca_", today, ".png", sep = ""), height = 7, width = 7)

# test if PC's correlate with selection coefficient
plot(pca_results$PC1, y)
cor.test(pca_results$PC1, y, method = "spearman")

plot(pca_results$PC2, y)
cor.test(pca_results$PC2, y, method = "spearman")

plot(pca_results$PC3, y)
cor.test(pca_results$PC3, y, method = "spearman")
```

# compare predictions to actual values on testing dataset
```{r}
#pred_vs_act = read.table("../workflow/predicted_vs_actual.txt")
train_pred_vs_act = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/selection-demography-cnn/workflow/train_predicted_vs_actual.txt")

val_pred_vs_act = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/selection-demography-cnn/workflow/val_predicted_vs_actual.txt")

test_pred_vs_act = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/selection-demography-cnn/workflow/test_predicted_vs_actual.txt")

# subset data to remove outliers
#pred_vs_act = pred_vs_act[(pred_vs_act$V1 <= 5000),]

# distribution of fixation times

## training set
train_pred_vs_act_melt = melt(train_pred_vs_act, id = c("V1", "V4"))
train_pred_vs_act_melt$variable = as.character(train_pred_vs_act_melt$variable)
train_pred_vs_act_melt$variable[(train_pred_vs_act_melt$variable == "V2")] = "Actual"
train_pred_vs_act_melt$variable[(train_pred_vs_act_melt$variable == "V3")] = "Predicted"

ggplot(train_pred_vs_act_melt, aes(x = value, color = variable)) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(Fixation time)", color = "Type", title = "Training set")

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/selection-demography-cnn/results/", today, "/train_actual-vs-predicted_density_", today, ".png", sep = ""), height = 7, width = 7)

## validation set
#val_pred_vs_act$V2 = log10(val_pred_vs_act$V2)
val_pred_vs_act_melt = melt(val_pred_vs_act, id = c("V1", "V4"))
val_pred_vs_act_melt$variable = as.character(val_pred_vs_act_melt$variable)
val_pred_vs_act_melt$variable[(val_pred_vs_act_melt$variable == "V2")] = "Actual"
val_pred_vs_act_melt$variable[(val_pred_vs_act_melt$variable == "V3")] = "Predicted"

ggplot(val_pred_vs_act_melt, aes(x = value, color = variable)) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(Fixation time)", color = "Type", title = "Valdiation set")

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/selection-demography-cnn/results/", today, "/val_actual-vs-predicted_density_", today, ".png", sep = ""), height = 7, width = 7)

## testing set
#test_pred_vs_act$V2 = log10(test_pred_vs_act$V2)
test_pred_vs_act_melt = melt(test_pred_vs_act, id = c("V1", "V4"))
test_pred_vs_act_melt$variable = as.character(test_pred_vs_act_melt$variable)
test_pred_vs_act_melt$variable[(test_pred_vs_act_melt$variable == "V2")] = "Actual"
test_pred_vs_act_melt$variable[(test_pred_vs_act_melt$variable == "V3")] = "Predicted"

ggplot(test_pred_vs_act_melt, aes(x = value, color = variable)) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(Fixation time)", color = "Type", title = "Testing set")

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/selection-demography-cnn/results/", today, "/test_actual-vs-predicted_density_", today, ".png", sep = ""), height = 7, width = 7)

# actual vs predicted values

##training set
my_cor_coeff = cor.test(train_pred_vs_act$V2, train_pred_vs_act$V3, method = "pearson")

cor_est = signif(my_cor_coeff$estimate, digits = 3)
cor_lwb = signif(my_cor_coeff$conf.int[1], digits = 3)
cor_upb = signif(my_cor_coeff$conf.int[2], digits = 3)
cor_pvalue = signif(my_cor_coeff$p.value, digits = 3)

ggplot(aes(x = V3, y = V2), data = train_pred_vs_act) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "#8C0172", linewidth = 2) +
  geom_vline(xintercept = mean(train_pred_vs_act$V2), linewidth = 1, linetype = "dotted", color = "red") + 
  geom_hline(yintercept = mean(train_pred_vs_act$V3), linewidth = 1, linetype = "dotted", color = "red") + 
  geom_smooth(method = "lm", color = "#9B951B", linewidth = 2) +
  #geom_smooth(method = "loess", color = "#B2F2FD", linewidth = 2) +
  theme_classic() +
  labs(y = "log10(Actual Fixation time)", 
       x = "log10(Predicted Fixation time)",
       title = paste("Training set: r = ", cor_est, ", 95 % CI = [",  cor_lwb, ", ", cor_upb, "] , p = ", cor_pvalue, sep = ""))

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/selection-demography-cnn/results/", today, "/train_actual-vs-predicted_scatter_", today, ".png", sep = ""), height = 7, width = 7)

## validation set
my_cor_coeff = cor.test(val_pred_vs_act$V2, val_pred_vs_act$V3, method = "pearson")

cor_est = signif(my_cor_coeff$estimate, digits = 3)
cor_lwb = signif(my_cor_coeff$conf.int[1], digits = 3)
cor_upb = signif(my_cor_coeff$conf.int[2], digits = 3)
cor_pvalue = signif(my_cor_coeff$p.value, digits = 3)

ggplot(aes(x = V2, y = V3), data = val_pred_vs_act) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "#8C0172", linewidth =
                2) +
  geom_vline(xintercept = mean(val_pred_vs_act$V2), linewidth = 1, linetype = "dotted", color = "red") + 
  geom_hline(yintercept = mean(val_pred_vs_act$V3), linewidth = 1, linetype = "dotted", color = "red") + 
  geom_smooth(method = "lm", color = "#9B951B", linewidth = 2) +
  #geom_smooth(method = "loess", color = "#B2F2FD", linewidth = 2) +
  theme_classic() +
  labs(x = "log10(Actual Fixation time)", 
       y = "log10(Predicted Fixation time)",
       title = paste("Validation set: r = ", cor_est, ", 95 % CI = [",  cor_lwb, ", ", cor_upb, "] , p = ", cor_pvalue, sep = ""))

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/selection-demography-cnn/results/", today, "/val_actual-vs-predicted_scatter_", today, ".png", sep = ""), height = 7, width = 7)

## test set
my_cor_coeff = cor.test(test_pred_vs_act$V2, test_pred_vs_act$V3, method = "pearson")

cor_est = signif(my_cor_coeff$estimate, digits = 3)
cor_lwb = signif(my_cor_coeff$conf.int[1], digits = 3)
cor_upb = signif(my_cor_coeff$conf.int[2], digits = 3)
cor_pvalue = signif(my_cor_coeff$p.value, digits = 3)

ggplot(aes(x = V3, y = V2), data = test_pred_vs_act) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "#8C0172", linewidth = 2) +
  geom_vline(xintercept = mean(test_pred_vs_act$V3), linewidth = 1, linetype = "dotted", color = "red") +
  geom_hline(yintercept = mean(test_pred_vs_act$V2), linewidth = 1, linetype = "dotted", color = "red") + 
  geom_smooth(method = "lm", color = "#9B951B", linewidth = 2) +
  #geom_smooth(method = "loess", color = "#B2F2FD", linewidth = 2) +
  theme_classic() +
  labs(y = "log10(Actual Fixation time)", 
       x = "log10(Predicted Fixation time)",
       title = paste("Testing set: r = ", cor_est, ", 95 % CI = [",  cor_lwb, ", ", cor_upb, "] , p = ", cor_pvalue, sep = ""))

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/selection-demography-cnn/results/", today, "/test_actual-vs-predicted_scatter_", today, ".png", sep = ""), height = 7, width = 7)

# standard deviation of predictions 

## test set
test_cor_std_vs_act = cor.test(test_pred_vs_act$V2, test_pred_vs_act$V4, method = "pearson")

cor_est = signif(test_cor_std_vs_act$estimate, digits = 3)
cor_lwb = signif(test_cor_std_vs_act$conf.int[1], digits = 3)
cor_upb = signif(test_cor_std_vs_act$conf.int[2], digits = 3)
cor_pvalue = signif(test_cor_std_vs_act$p.value, digits = 3)

ggplot(aes(x = V2, y = V4), data = test_pred_vs_act) +
  geom_point() +
  geom_vline(xintercept = mean(test_pred_vs_act$V2), linewidth = 1, linetype = "dotted", color = "red") + 
  geom_hline(yintercept = mean(test_pred_vs_act$V4), linewidth = 1, linetype = "dotted", color = "red") + 
  geom_smooth(method = "lm", color = "#9B951B", linewidth = 2) +
  #geom_smooth(method = "loess", color = "#B2F2FD", linewidth = 2) +
  theme_classic() +
  labs(x = "log10(Actual Fixation time)", 
       y = "SD in log10(Predicted Fixation time)",
       title = paste("Testing set: r = ", cor_est, ", 95 % CI = [",  cor_lwb, ", ", cor_upb, "] , p = ", cor_pvalue, sep = ""))

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/selection-demography-cnn/results/", today, "/test_actual-vs-predicted-sd_scatter_", today, ".png", sep = ""), height = 7, width = 7)

## validation set
val_cor_std_vs_act = cor.test(val_pred_vs_act$V2, val_pred_vs_act$V4, method = "pearson")

cor_est = signif(val_cor_std_vs_act$estimate, digits = 3)
cor_lwb = signif(val_cor_std_vs_act$conf.int[1], digits = 3)
cor_upb = signif(val_cor_std_vs_act$conf.int[2], digits = 3)
cor_pvalue = signif(val_cor_std_vs_act$p.value, digits = 3)

ggplot(aes(x = V2, y = V4), data = val_pred_vs_act) +
  geom_point() +
  geom_vline(xintercept = mean(val_pred_vs_act$V2), linewidth = 1, linetype = "dotted", color = "red") + 
  geom_hline(yintercept = mean(val_pred_vs_act$V4), linewidth = 1, linetype = "dotted", color = "red") + 
  geom_smooth(method = "lm", color = "#9B951B", linewidth = 2) +
  #geom_smooth(method = "loess", color = "#B2F2FD", linewidth = 2) +
  theme_classic() +
  labs(x = "log10(Actual Fixation time)", 
       y = "SD in log10(Predicted Fixation time)",
       title = paste("Validation set: r = ", cor_est, ", 95 % CI = [",  cor_lwb, ", ", cor_upb, "] , p = ", cor_pvalue, sep = ""))

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/selection-demography-cnn/results/", today, "/val_actual-vs-predicted-sd_scatter_", today, ".png", sep = ""), height = 7, width = 7)

# variation in predictions vs sweep age
# ggplot(aes(x = V2, y = V4), data = val_pred_vs_act) +
#   geom_point() +
#   geom_vline(xintercept = mean(val_pred_vs_act$V2), linewidth = 1, linetype = "dotted", color = "red") + 
#   geom_hline(yintercept = mean(val_pred_vs_act$V4), linewidth = 1, linetype = "dotted", color = "red") + 
#   geom_smooth(method = "lm", color = "#9B951B", linewidth = 2) +
#   #geom_smooth(method = "loess", color = "#B2F2FD", linewidth = 2) +
#   theme_classic() +
#   labs(x = "log10(Actual Fixation time)", 
#        y = "SD in log10(Predicted Fixation time)",
#        title = paste("Validation set: r = ", cor_est, ", 95 % CI = [",  cor_lwb, ", ", cor_upb, "] , p = ", cor_pvalue, sep = ""))

# actual vs coefficient of variation in predictions
ggplot(aes(x = V2, y = V4/V3), data = test_pred_vs_act) +
  geom_point() +
  geom_smooth(method = "lm", color = "#9B951B", linewidth = 2) +
  #geom_smooth(method = "loess", color = "#B2F2FD", linewidth = 2) +
  theme_classic() +
  labs(x = "log10(Actual Fixation time)", 
       y = "CV in log10(Predicted Fixation time)",
       )

# Measure error between actual and predicted
# How does error correlate with simulation parameters?
test_pred_vs_act$error = (test_pred_vs_act$V2 - test_pred_vs_act$V3)^2
names(test_pred_vs_act)[1] = "ID"
errors = merge(params, test_pred_vs_act, by = "ID")

errors$NQ = errors$N/errors$Q
errors$SQ = errors$sweepS*errors$Q
errors$MQ = errors$mu*errors$Q
errors$RQ = errors$R*errors$Q
errors$TQ = errors$tau*errors$Q

error_mod = lm(log10(error) ~ NQ + log10(SQ) + sigma + h + sigma*h + log10(MQ) + log10(RQ) + TQ + B + U, data = errors)
summary(error_mod)
plot(error_mod)

step(error_mod)

error_mod = lm(log10(error) ~ log10(SQ) + NQ + h, data = errors)
summary(error_mod)
```
# Adjust model predictions so that they have a 1:1 relationship with simulations
```{r}

mod_pred_vs_act = lm(V3 ~ V2, data = test_pred_vs_act)
coef_pred_vs_act = coefficients(mod_pred_vs_act)
res_pred_vs_act = residuals(mod_pred_vs_act)

#y_star = test_pred_vs_act$V3 + (1 - coef_pred_vs_act[2])*test_pred_vs_act$V2 - coef_pred_vs_act[1]
y_star = (test_pred_vs_act$V3 - coef_pred_vs_act[1])/coef_pred_vs_act[2]

plotdata = data.frame(
  y_star = y_star,
  x = test_pred_vs_act$V2
)

#
ggplot(aes(x = x, y = y_star), data = plotdata) +
  geom_point() +
  #geom_smooth(method = "lm", color = "#9B951B", linewidth = 2) +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  labs(x = "log10(Actual Fixation time)", 
       y = "log10(Predicted Fixation time, adjusted)")

cor.test(plotdata$x, plotdata$y_star)

# plot histogram
plotdata_melt = melt(plotdata)
plotdata_melt$variable = as.character(plotdata_melt$variable)
test_pred_vs_act_melt$variable[(plotdata_melt$variable == "V2")] = "Actual"
test_pred_vs_act_melt$variable[(plotdata_melt$variable == "V3")] = "Predicted"

ggplot(test_pred_vs_act_melt, aes(x = value, color = variable)) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(Fixation time)", color = "Type", title = "Testing set")
```


# How does time to fixation in simulations vary with the simulation parameters?
```{r}
#fix_times_files = list.files(path = "../workflow/data/fix_times", full.names = T)
# fix_times_files = list.files(path = "/mnt/scratch/robe1195/Josephs_Lab_Projects/selection-demography-cnn/workflow/data/fix_times", full.names = T)
# fix_times = lapply(fix_times_files, fread)

#fail_files = list.files(path = "../workflow/data/fails", full.names = T)
#fail_files = list.files(path = "/mnt/scratch/robe1195/Josephs_Lab_Projects/selection-demography-cnn/workflow/data/fails", full.names = T)
#fail_times = lapply(fail_files, fread)

# extract_number = function(x){
#   x$V1[1]
# }
# 
# fix_times = unlist(lapply(fix_times, extract_number))

#fail_times = unlist(lapply(fail_times, extract_number))

# fix_times = data.frame(
#   tf = fix_times,
#   ID = gsub(".txt", "", gsub(".*fix_time_", "",fix_times_files))
# )

#fail_times = data.frame(
#  fails = fail_times,
#  ID = gsub(".txt", "", gsub(".*fails_", "", fail_files))
#)

#fix_times = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/selection-demography-cnn/workflow/fixation_times.txt", col.names = c("ID", "tf"))

#fix_times = merge(params, fix_times, by = "ID")
#fix_times = merge(fix_times, fail_times, by = "ID")

# convert tf to units of Ne
#fix_times$tfne = fix_times$tf/fix_times$Ne

# square selfing rate
#fix_times$sigma2 = fix_times$sigma^2

# remove any simulations that don't have images
image_ids = list.files(path = "/mnt/scratch/robe1195/Josephs_Lab_Projects/selection-demography-cnn/workflow/data/images", full.names = T)

image_ids = as.numeric(gsub(".*_", "", gsub(".png", "", image_ids)))

params = params[(params$ID %in% image_ids),]

# calculate probability of failure
# For each simulation, what geometric distribution would give the observed number of failures as it's expected value
#fix_times$fix_prob = 1/(fix_times$fails + 1)

# exploratory analysis, look at distribution of fix times
ggplot(fix_times, aes(x=tf)) +
  geom_density() +
  theme_classic() +
  labs(x = "fixation times")

ggplot(fix_times, aes(x=fails)) +
  geom_density() +
  theme_classic() +
  labs(x = "failures")

ggplot(fix_times, aes(x=log10(tf))) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(fixation times)")

ggplot(fix_times, aes(x = log10(sweepS*Q), y = log10(tf))) +
  geom_point() +
  theme_classic()

ggplot(fix_times, aes(x=log10(fails + 1))) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(failures + 1)")

ggplot(fix_times, aes(x=fix_prob)) +
  geom_density() +
  theme_classic() +
  labs(x = "Probability of fixation")

ggplot(fix_times, aes(x=log10(tf), y = log10(fails + 1))) +
  geom_point() +
  geom_smooth() +
  theme_classic() +
  labs(x = "Fixation time (generations)", y = "Failures")

cor.test(fix_times$tf, fix_times$fails, method = "spearman")

# change growth rate (r) to categorical variable
fix_times$rcat = NA
fix_times$rcat[(fix_times$r == 0)] = "constant"
fix_times$rcat[(fix_times$r > 0) & (fix_times$r < 0.5) & (fix_times$N < fix_times$K)] = "growing"
fix_times$rcat[(fix_times$r > 0) & (fix_times$r < 0.5) & (fix_times$N > fix_times$K)] = "shrinking"
fix_times$rcat[(fix_times$r > 2) & (fix_times$r < sqrt(6))] = "2-cycle"
fix_times$rcat[(fix_times$r > sqrt(6))] = "chaos"

table(fix_times$rcat)

fix_times$rcat = relevel(as.factor(fix_times$rcat), ref = "constant")

# calculate Ne
ggplot(fix_times[(fix_times$sweepS < 1/(fix_times$N*(2 - fix_times$sigma)/2)) & fix_times$rcat == "constant",], aes(x = N*(2-sigma)/2, y = tf)) +
  geom_point() +
  geom_abline(slope = 4, intercept = 0) +
  theme_classic()

# try linear models, see if there are any problems
tf_mod = lm(log10(tf) ~ log10(sweepS*Q) + h + log10(mu*Q) + log10(R*Q) + tau/Q + sigma + sigma*h + f0 + f1 + N/Q + n + lambda + n*lambda, data = fix_times)
summary(tf_mod)
plot(tf_mod)

fail_mod = lm(log10(fails + 1) ~ log10(sweepS) + h + log10(mu) + log10(R) + tau + sigma + sigma2 + sigma*h + f0 + f1 + N + n + rcat, data = fix_times)
summary(fail_mod)
plot(fail_mod)

# try glm instead of lm for fixation time
#tf_mod = glm(tf ~ log10(sweepS)+ h + log10(mu) + log10(R) + tau + f0 + f1 + N + lambda + n + r + sigma + sigma2 + sigma*h, data = fix_times, family = Gamma(link = "log"))
#summary(tf_mod)
#plot(tf_mod)

# fit model for failures
fail_mod = glm(fails ~ log10(sweepS)+ h + log10(mu) + log10(R) + tau + sigma + sigma2 + sigma*h + sigma*log10(R) + f0 + f1 + N + lambda + n + n*lambda + rcat, data = fix_times, family = "quasipoisson")
summary(fail_mod)
plot(fail_mod)

# find model with fewest terms but still lots of power
step(tf_mod, direction = "both")

# parametric plot of fixation time versus other parameters
ggplot(fix_times, aes(sigma, h, color = log(tf))) +
  geom_point() +
  theme_classic()

# re-fit model for only recessive mutations
tf_mod = lm(log10(tf) ~ log10(sweepS) + h + log10(mu) + log10(R) + sigma + sigma2 + sigma*h + sigma*log10(R) + N + n, data = fix_times[(fix_times$h < 0.5 & fix_times$f0 == 0 & fix_times$f1 == 1 & fix_times$rcat == "constant" & fix_times$M < 1),])
summary(tf_mod)

cor.test(fix_times$tf[(fix_times$h < 0.5)], fix_times$sigma[(fix_times$h < 0.5)])
```

# Stratified sampling along time to fixation
```{r}
# split
#widths_to_try = seq(from = 0.01, to = 1, by = 0.05)
#count_thresh = 1000

#for(width in widths_to_try){
#  breaks = seq(from = min(log10(fix_times$tf)), to = max(log10(fix_times$tf)), by = width)
#  fix_times$bin = cut(log10(fix_times$tf), breaks = breaks)
#  kept_bins = table(fix_times$bin)[table(fix_times$bin) > count_thresh]
#  total_in_downsample = min(kept_bins)*length(kept_bins)
#  print(paste("width: ", width, "; total: ", total_in_downsample))
#}

# focus on just hard sweeps
#fix_times = fix_times[(fix_times$f0 == 0 & fix_times$f1 == 1),]

width = 0.05
max_bin_height = 386

breaks = seq(from = min(log10(params$tf)) - width, to = max(log10(params$tf)) + width, by = width)
params$bin = cut(log10(params$tf), breaks = breaks)

table(params$bin)
max_bin_height = min(table(params$bin)[table(params$bin) >= max_bin_height])

strat_sample = NULL
for(bin in unique(params$bin)){
  fix_times_bin = params[(params$bin == bin),]
  if(nrow(fix_times_bin) >= max_bin_height){
    down_sampled_bin = fix_times_bin[sample(1:nrow(fix_times_bin), replace = F, size = max_bin_height),]
    strat_sample = rbind(strat_sample, down_sampled_bin)
  }
}

table(strat_sample$bin)
 
# check that tf is now uniformly distributed across data
ggplot(strat_sample, aes(x = log10(tf))) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(Fixation time)")

ggplot(strat_sample, aes(x = log10(tf))) +
  geom_boxplot() +
  theme_classic() +
  labs(x = "log10(Fixation time)")

ggplot(strat_sample, aes(x = N)) +
  geom_density() +
  theme_classic() +
  labs(x = "Population size")

ggplot(strat_sample, aes(x = log10(mu))) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(mutation rate)")

ggplot(strat_sample, aes(x = log10(R))) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(recombination rate)")

ggplot(strat_sample, aes(x = tau)) +
  geom_density() +
  theme_classic() +
  labs(x = "Generations post-fixation when observed")

ggplot(strat_sample, aes(x = sigma)) +
  geom_density() +
  theme_classic() +
  labs(x = "Selfing rate")

ggplot(strat_sample, aes(x = h)) +
  geom_density() +
  theme_classic() +
  labs(x = "dominance coefficient")

ggplot(strat_sample, aes(x = log10(sweepS))) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(selection coefficient)")

ggplot(fix_times, aes(x = log10(sweepS))) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(selection coefficient)")

ggplot(strat_sample, aes(x = f0)) +
  geom_density() +
  theme_classic() +
  labs(x = "neutral -> benficial frequency")

ggplot(strat_sample, aes(x = f1)) +
  geom_density() +
  theme_classic() +
  labs(x = "beneficial -> neutral frequency")

ggplot(strat_sample, aes(x = B)) +
  geom_density() +
  theme_classic() +
  labs(x = "proportion of linked mutations which are beneficial")

ggplot(strat_sample, aes(x = U)) +
  geom_density() +
  theme_classic() +
  labs(x = "proportion of linked mutations which are deleterious")

ggplot(strat_sample, aes(x = M)) +
  geom_density() +
  theme_classic() +
  labs(x = "proportion of linked mutations which are neutral")

# Look at correlation between selection coefficient and tf
ggplot(strat_sample, aes(x = log10(sweepS*Q), y = log10(tf))) +
  geom_point() +
  theme_classic()

# randomly subsample data into training, testing, and validation
train_n = round(nrow(strat_sample)*0.8)
test_n = round(nrow(strat_sample)*0.1)
val_n = round(nrow(strat_sample)*0.1)

excess = (train_n + test_n + val_n) - nrow(strat_sample) # remove excess sims from partitioning
train_n = train_n - excess

train_n + test_n + val_n == nrow(strat_sample) # output should be TRUE

split_column = c(rep("train", times = train_n), rep("test", times = test_n), rep("val", times = val_n))

strat_sample$split = sample(split_column, replace = F, size = nrow(strat_sample))

table(strat_sample$split)

# What mean squared error would you expect if model predicted the mean for every image?
# This naive model would result in a loss equal to the variance of the response
var(log10(strat_sample[(strat_sample$split == "val"), "tf"]))

# check that dependent variable is still uniform after stratification
ggplot(strat_sample[(strat_sample$split == "train"),], aes(x=log10(tf))) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(fixation times)")

ggplot(strat_sample[(strat_sample$split == "test"),], aes(x=log10(tf))) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(fixation times)")

ggplot(strat_sample[(strat_sample$split == "val"),], aes(x=log10(tf))) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(fixation times)")

# save stratified sample
write.table(strat_sample, "stratified_sample_12.tsv", sep = "\t", quote = F, row.names = F)
```

# subset simulations that are outside range of training data
```{r}
low_examples = params[(log10(params$tf) < min(train_pred_vs_act$V2)),]

high_examples = params[(log10(params$tf) > max(train_pred_vs_act$V2)),]

```

# evaluate results from real images
```{r}
real_pred = read.table("/mnt/scratch/robe1195/vcf_to_images/real_predictions.txt", header = T)

ggplot(real_pred, aes(x=tf_mean)) +
  geom_density() +
  theme_classic() +
  labs(x = "log10(predicted fixation times)")

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/selection-demography-cnn/results/", today, "/real_density_10missing_", today, ".png", sep = ""), height = 7, width = 7)

# parse figure ids
split_ids = strsplit(real_pred$ID, "_")

extract_data = function(x, i){
  x[i]
}

real_pred$chrom = unlist(lapply(split_ids, FUN = extract_data, i = 1))
real_pred$start = unlist(lapply(split_ids, FUN = extract_data, i = 2))
real_pred$end = unlist(lapply(split_ids, FUN = extract_data, i = 3))

# plot 
ggplot(real_pred[(real_pred$chrom == "1"),], aes(x = start, y = tf_mean)) +
  geom_point() +
  theme_classic() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  labs(x = "Position", y = "tf")

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/selection-demography-cnn/results/", today, "/real_manhattan_chrom1_", today, ".png", sep = ""), height = 7, width = 7)
```


#
```{r}
expected_seg_sites = function(N,mu,sigma,L){

4*mu*(N/(1 + sigma/(2-sigma)))*L

}

expected_seg_sites(1000,1e-8,0,1e6)
```

#
```{r}
s = runif(6000)
h = runif(6000)

fixtimes = data.frame(
  id = 1:6000,
  hs = h*s
)

fixtimes$bins = cut(fixtimes$hs, breaks = 10)

target = min(table(fixtimes$bins))

for(bin in unique(fixtimes$bins)){
    fixsub = fixtimes[(fixtimes$bins == bin),]
    fixsub[(fixsub$id %in% sample(fixsub$id, size = target, replace = F)),]
}

table(bins)


```

# shape data into one matrix
```{r}
# get max number of variants so, we know what to pad to
rowPadMax = max(unlist(lapply(vcfs, nrow)))

# zero pad or subsample data as necessary so all tables have same dimmensions
zeropad = function(x, rowPadMax){
  varCount = nrow(x)

  # subsample data if there are too many variants
  if(varCount > rowPadMax){
    x = x[sort(sample(1:varCount, rowPadMax, replace = F)),]
  }

  # Add zero-padds if there are too few variants
  if(varCount < rowPadMax){
    lastVar = x[varCount,"POS"]
    for(i in 1:(rowPadMax - varCount)){
      x = rbind(x, c(1, lastVar + i, "MT=0;", rep(0, times = 128)))
    }
  }
  
  return(x)
}

vcfs = lapply(vcfs, zeropad, rowPadMax = rowPadMax)

# melt vcfs
meltvcfs = function(x){
  x = melt(x, id.vars = c("CHROM", "POS", "INFO"))
  return(x[,5])
}

vcfs = lapply(vcfs, meltvcfs)

# combine all vcfs into one matrix
vcfs = do.call("cbind", vcfs)
dim(vcfs)

# convert all columns to numeric
vcfs = apply(vcfs, MARGIN = 2, as.numeric)
dim(vcfs)
```

# PCA and correlations
```{r}
# simulations should be rows
tvcfs = t(vcfs)

# tvcfsNoZero = tvcfs[,apply(tvcfs, 2, function(x){length(unique(x)) > 1})]

# do PCA
vcfpca = prcomp(tvcfs, center = T)

plot(vcfpca)
plot(vcfpca$x)

# If outcome is binary, look for association using logistic regression
regframe = as.data.frame(cbind(y, vcfpca$x))
regframe$y = as.factor(regframe$y)

mod = glm(y ~ PC1 + PC2 + PC3 + PC4 + PC5 + PC6, data = regframe, family = "binomial")
summary(mod)

# visually show how PCs correlate with dependent variable
ggplot(regframe, aes(PC1, PC2, color = y)) +
  geom_point() +
  theme_classic() +
  scale_color_scico_d(palette = "hawaii", labels = c("Neutral", "Selective sweep"), name = "Simulation")

ggsave("pca.png", scale = 0.5, width = 10, height = 7)

# correlate PCs with the dependent variable
plot(vcfpca$x[,1], y)
cor.test(vcfpca$x[,1], y, method = "spearman")

plot(vcfpca$x[,2], log(y))
cor.test(vcfpca$x[,2], log(y), method = "spearman")
```

